---
title: "p8105_hw1_qc2385"
author: "Qixuan Cao"
date: "2026-09-26"
output: github_document
---

## Problem 1
Let's read the penguins dataset:
```{r load pegunis}
data("penguins", package = "palmerpenguins")
```

The penguins dataset contains measurements and characteristics of penguins.

It has `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The variables are `r paste(names(penguins), collapse = ", ")`.

The species are `r paste(unique(penguins$species), collapse = ", ")`,

The islands are `r paste(unique(penguins$island), collapse = ", ")`.

Bill length and flipper length are measured in millimeters.

The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 1)` mm.

### scatterplot
```{r}
library(ggplot2)
```
```{r peguin_plot}
scatter_plot = ggplot(
  penguins,
  aes(
    x = bill_length_mm, 
    y = flipper_length_mm, 
    color = species
      )
) + 
  geom_point()

scatter_plot
```

## Problem 2
```{r}
library(tidyverse)
```
Create the data frame
```{r hw1_df}
hw1_df = 
  tibble(
    norm_sample = rnorm(10, mean = 0, sd = 1),
    greater_than_0 = norm_sample > 0,
    char_var = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_var = factor(rep(c("low", "medium", "high"), length.out =10))
  )

hw1_df

```
Try to take the mean of each variable in the data frame
```{r mean}
mean(pull(hw1_df, norm_sample))
mean(pull(hw1_df, greater_than_0))
mean(pull(hw1_df, char_var))
mean(pull(hw1_df, factor_var))
```
The mean of `norm_sample` works because it is numeric. 

The mean of `greater_than_0` also works because R treats `TRUE` as 1 and `FALSE` as 0;

The means of `char_var` and `factor_var` return `NA` because these variables are not numeric or logical.

applies the as.numeric function:
```{r convert_to_numeric, eval=FALSE}
as.numeric(hw1_df$greater_than_0)
as.numeric(hw1_df$char_var)
as.numeric(hw1_df$factor_var)
```
Converting the logical variable to numeric changes `TRUE` to 1 and `FALSE` to 0. This explains why its mean gives the proportion of values greater than zero. 

The character values are letters, so they cannot be converted to numbers and become NA.

Converting the factor gives the numeric codes of its levels (`high` = 1, `low` = 2, and `medium` = 3). These codes are category labels, not meaningful numerical measurements, so calculating their mean would not be useful.

Therefore,`mean()` only works when a variable can be given a genuine numeric interpretation.

